Linear Algebra as a Unifying Theory\\of Wireless Workload Modeling

### 2008-02-02 paper.tex
Rewriting the packets-bytes-duration example paragraph.

\begin{abstract}
\input{abstract.tex}
\end{abstract}

@@

The traffic workload pattern in a network is simply a collection of sequences of packets between hosts in the network sent at certain times. It is standard in network analysis to aggregate sequences of packets sharing the same \caps{IP} protocol, source and destination nodes, and source and destination \caps{TCP/UDP} port numbers. Such a sequence of packets is called a \textit{flow}. The traffic pattern is characterized completely by the detailed behavior of it's flows. The behavior of a is defined by six properties: its source and destination nodes, its start and end times, and the sizes of and intervals between its constituent packets. We propose to represent the behavior of each flow in terms of six vectors---one for each aspect of the flow's behavior. The aggregate behavior of all the flows in a workload pattern may then be represented by six matrices, each row of which gives the behavior vectors of one of the flows.

When mapping flow behaviors into a vector space, it is important that the mapping of behaviors into the vector spaces in question respect the vector space operations: addition of vectors and scalar multiplication. This requirement is best clarified with a simple example. Suppose that the behavior of a flow is represented by the number of packets, $p$, the total bytes transmitted, $b$, and its total duration, $d$. This triplet can be naturally embedded into Euclidean space: $\vec{x}=(p,b,d)\in\R^3$. If we have two flows with behavior vectors $\vec{x}_1$ and $\vec{x}_2$, then their ``sum'' behavior, $\vec{x}_1+\vec{x}_2$ has $p_1+p_2$ packets, $b_1+b_2$ bytes, $d_1+d_2$ duration. Their average, i.e. $\frac{1}{2}(\vec{x}_1+\vec{x}_2)$, has the average number of packets, bytes, and duration, which is precisely what the ``average behavior'' of the two flows should intuitively be.

@@

To expand on this, $z_i$ (i.e. the $i$th element of $\vec{z}$) is the number of packets in the flow such with $i$ bytes and $d_z$ is the largest possible packet size, which in practice is the maximum transfer unit of the wireless medium---typically 1500 bytes. The histogram of inter-packet intervals is similar, except that domain of possible intervals is a continuum, so we much map the range of possible intervals onto the discrete index values $\left\{1,\dots,d_v\right\}$. In our experiments, we use the following logarithmic quantization function:
\begin{align}
q(v) &= \left\lfloor(d_v-1)\frac{\ln v - \ln v_{\min}}{\ln v_{\max}-\ln v_{\min}}\right\rfloor+1.
\end{align}
with $v_{\min}=10^{-6}$, which is the minimal interval our trace data, and $v_{\max}=600$, which is the total duration of each simulated situation, and hence an upper bound on the inter-packet intervals. We use $d_v=1500$ to match the dimensionally of the packet size representation. The packet size and inter-packet intervals are collected into matrices, $Z\in\R^{f \times d_z}$ and $V\in\R^{f \times d_v}$, whose rows are the behavior vectors $\vec{z}_i$ and $\vec{v}_i$ indexed from $i=1,\dots,f$ where $f$ is the number of flows in the collection.

>>>>

%\begin{abstract}
%\input{abstract.tex}
%\end{abstract}

@@

The traffic workload pattern in a network is simply a collection of packets between hosts in the network sent at certain times. It is standard in network analysis, however, to aggregate sequences of packets sharing the same \caps{IP} protocol, source and destination nodes, and source and destination \caps{TCP} or \caps{UDP} port numbers. Such a sequence of packets is called a \textit{flow}. The traffic pattern is characterized completely by the detailed behavior of it's flows. The behavior of a flow is defined by six properties: its source and destination nodes, its start and end times, and the specific sizes of and intervals between its constituent packets. We propose to represent the behavior of each flow in terms of six vectors---one for each aspect of its behavior. The aggregate behavior of all the flows in a workload pattern may then be represented by six matrices, the rows of which are the behavior vectors of the flows.

When mapping flow behaviors into a vector space, it is important that the mapping of behaviors into the vector spaces in question respect the vector space operations: addition of vectors and scalar multiplication. This requirement is best clarified with a simplified example. Suppose that the behavior of a flow is represented by the number of packets, $p$, the total bytes transmitted, $b$, and its total duration, $d$. Despite its simplicity, this representation is adequate to reproduce constant bit-rate (\caps{CBR}), time-randomized flow behavior. The triplet of representative values can be naturally embedded into Euclidean space: $\vec{x}=(p,b,d)\in\R^3$. If we have two flows with behavior vectors $\vec{x}_1$ and $\vec{x}_2$, then their sum as vectors, $x_1+x_2=(p_1+p_2,b_1+b_2,d_1+d_2)$, provides an appropriate description of the aggregate behavior of the two flows together. Similarly, their average, i.e. $\frac{1}{2}(\vec{x}_1+\vec{x}_2)$, has the average number of packets, bytes, and duration, which is precisely what the ``average behavior'' of the two flows should intuitively be.

@@

The packets size behavior of a flow is represented by a histogram vector, $\vec{z}$, whose $i$th element, $z_i$, is the number of packets in the flow with $i$ bytes of payload data. The dimension, $d_z$, is the largest possible packet size, which in practice is the maximum transfer unit of the wireless medium---typically 1500 bytes. The histogram of inter-packet intervals is similar, except that domain of possible intervals is a continuum, so we much map the range of possible intervals onto the discrete index values $\left\{1,\dots,d_v\right\}$. In our experiments, we use the following logarithmic quantization function:
\begin{align}
q(v) &= \left\lfloor(d_v-1)\frac{\ln v - \ln v_{\min}}{\ln v_{\max}-\ln v_{\min}}\right\rfloor+1.
\end{align}
with $v_{\min}=10^{-6}$, which is the minimal interval our trace data, and $v_{\max}=600$, which is the total duration of each simulated situation, and hence an upper bound on the inter-packet intervals. We use $d_v=1500$ to match the dimensionally of the packet size representation. The packet size and inter-packet intervals are collected into matrices, $Z\in\R^{f \times d_z}$ and $V\in\R^{f \times d_v}$, whose rows are the behavior vectors $\vec{z}_i$ and $\vec{v}_i$ indexed from $i=1,\dots,f$ where $f$ is the number of flows in the collection.
